Same time last year, I was completely lost.
It was only because of the people who came across my path that this journey has come to an end.

\paragraph{}
This is the minimum tribute I could pay to them; friends, colleagues and advisors towards who I feel only sincere gratitude and will always respect.

\paragraph{}
Foremost, I owe a very important debt to Dr. Sirivianos Michael, advisor and mentor of this thesis and many other aspects of life.
I am thankful for your patience and unrestrictive guidance.
It is fulfilling to find a person you can look up to.

\paragraph{}
Likewise, I truly appreciate the feedback offered by Erik Nordstrom from Princeton SNS Lab, when encountering problems with the internals of the Serval architecture.

\paragraph{}
I would also like to particularly thank Bernard Van De Walle, Product Manager at Nuage Networks, and Erik Neel, head IPD QA Antwerp at Alcatel-Lucent, for their insightful feedback and for helping me realize the huge potential of network virtualization in datacenters.
Service-Aware networking is in a great level a brainchild of your advice.

\paragraph{}
Furthermore, I am grateful to Kosiaris Alex and the team of grnet (Greek Research \& Technology Network), for willingly offering okeanos cloud service as an experiments testbench.

\paragraph{}
Moreover, I have received generous support from Apache Cloudstack developers, as well as by "anonymous" contributors of the Open Source movement in general, who have created such amazing projects as Serval, nginx, Linux, perf, git and so many others.

\paragraph{}
My intellectual debt is to Korina Patelis, fellow hacktivist and Internet pioneer, since her intuition that users will not always want to connect to the big-mother Internet, inspired a whole new chapter on the DHT-based resolution of service names.\\
\indent For the same reason to Komaitis Konstantinos, Policy Advisor at the Internet Society, for his instructive publications and for accepting to discuss on service resolution approaches in regard to today's DNS.

\paragraph{}
The stimulating discussions with Papadopoulos Fragkiskos, Mathieu Goessens, Marios Fotiou and others have broaden the horizons of this thesis.

\paragraph{}
It's time to thank my friends who have been close this whole time, among them Achilleas, Avgoustis, Christos, Chrystalleni, Desi, George, Georgia, Giannis, Kostas, Maria, Marialena, Michalis, Nicholas, Nick, Olia, Rodoula, Vana and the Space Apps and \#hack66 peers.

\paragraph{}
Nevertheless, I mean the deepest gratitude to my parents, Parthenios and Georgia, for their endless love and support. And of course to my dear sister Dimitra, with whom we will be graduating the same day.\\[0.8cm]


\begin{flushright}
\emph{I would like to express my heartfelt appreciation to all of you,\\
without your constant encouragement this dissertation\\
would not have been possible.}
\end{flushright}